---
title: "Community Analysis"
output: 
  html_document:
    toc: true
    toc_depth: 4
---

<link rel="stylesheet" type="text/css" href="http://cdn.datatables.net/1.10.5/css/jquery.dataTables.min.css">
<script src="http://code.jquery.com/jquery-2.1.2.min.js"></script>
<script src="http://cdn.datatables.net/1.10.5/js/jquery.dataTables.min.js"></script>

<script type="text/javascript">
         $(document).ready(function() {
             $(".dtable").DataTable();
         } );
</script>


```{r init, echo = FALSE, warning=FALSE, message=FALSE}
here::i_am('docs/community-analysis.Rmd')
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  echo = FALSE
)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message= FALSE)

# Define a custom knitr hook for collapsing output
knitr::knit_hooks$set(fig_collapse=function(before, options, envir) {
  if (before) {
    # Opening HTML tags for the collapsible container
    paste('<details><summary>Click to view figures</summary><div>')
  } else {
    # Closing HTML tags
    paste('</div></details>')
  }
})
## load the libraries needed
library(here)
library(junkR)
library(RODBC)
library(magrittr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(rstan)
library(purrr)
library(viridis)
update_data = FALSE
theme_set(theme_minimal())
# source the code to initiate the 
conn <<- odbcConnectAccess2007(here("ignore/Nearshore Survey.accdb"))

# source all the functions to read, clean, and update data products
purrr::walk(list.files(here::here('code/source'), pattern = ".*.R", full.names = TRUE), function(x) source(x))


```

Introduction
------

```{r load-biotow-data}

get_trawl_data(conn = conn, update = update_data)

```

## Data summaries:

How many 'bad' coordinates do we have? Either missing or with minute coordinates greater than 60.

`r bioTowTab %>% dplyr::filter(when_any(is.na(latStart_dd), is.na(lonStart_dd), is.na(latEnd_dd), is.na(lonEnd_dd), latme > 60)) %>% select(cno) %>% unique %>% count`: `r bioTowTab %>% dplyr::filter(when_any(is.na(latStart_dd), is.na(lonStart_dd), is.na(latEnd_dd), is.na(lonEnd_dd), latme > 60)) %>% select(cno) %>% unique %>% knitr::kable(.)`

How many missing weights?

`r bioTowTab %>% dplyr::filter(is.na(weight)) %>% count`

Missing weights with no corresponding length measurements?

`r bioTowTab %>% dplyr::filter(when_all(is.na(weight), is.na(flength), is.na(tlength))) %>% count`

## Length-mass relationships

While directly measured weights are present for most species, some of the data are censored due the precision of the balance, so weights bottom out at 0.01 and, for small species, exhibit distinct 'steps'. This creates an artificial truncation of the lower bound of the data and can also influence biomass estimates. This is likely not a huge deal for the purposes here, but if we want to incorporate this into any analyses, we can model masses and their uncertainty by incorporating the censored data structure directly into a length-mass model.

 
```{r censored-mass-ex}
spList = bioTowTab %>% 
  dplyr::filter(!when_all(is.na(tlength), is.na(flength), is.na(weight))) %>% 
  named_group_split(spn)

spnLWList = spList %>% map(~.x %>% 
    select(tlength, weight) %>% 
    na.omit %>% 
    # create a vector of censored measurements
    mutate(censored = ifelse(weight == 0.01, TRUE,FALSE))) %>% 
    # remove species that have fewer than 10 observations
    keep(~nrow(.) > 10)
# also remove American Lobster as they have a clear bi-modal L-M relationship
    # probably due to sex-specific differences
spnLWList['American Lobster'] <- NULL
    
# using Alewife as an example
spnLWList['Alewife'] %>% 
  map(~.x %>% 
        ggplot()+
        geom_point(aes(x = tlength, y = weight, color = censored))+
        scale_color_viridis(discrete = TRUE, end = 0.8)+
        scale_y_continuous(name = "mass (g)", limits = c(0, NA)))

```

```{r length-mass, warning=FALSE, message = FALSE}
spNames = names(spnLWList)

# debugonce(build_LM)
spNames %>% walk(~build_LM(specName = ., df = bioTowTab))

```

```{r plot length-mass-PPC, fig_collapse = TRUE}
lmFiles = list.files(here('ignore/models/'), pattern = '.*LM_censored.rds', full.names = TRUE)

lmNames = gsub(".*(/\\w+( \\w+)*_LM_censored)\\.rds","\\1", lmFiles) %>% 
  map_chr(~gsub("/(\\w+( \\w+)*)_LM_censored","\\1", .x))

lmFiles= setNames(lmFiles, nm = lmNames)

# debugonce(plot_LM_PPC)
map2(lmFiles, spnLWList, ~plot_LM_PPC(model_path = .x, df = .y))

```

## Collecting environmental variables

Here are the sampling dates from the trawl data.

<details>

```{r trawl-date-info}

bioTowTab %>% 
  mutate(towmonth = as.Date(paste0(lubridate::year(towDate),"-",lubridate::month(towDate),"-01"), format = "%Y-%m-%d")) %>%
  select(towmonth) %>% 
  distinct %>% 
  knitr::kable(.)
```

</details>

We can also view the spatial extent of the trawl data to confirm that they are within the bounds of the environmental variables.

```{r trawl-lines}
# create a metadata with the start and end of each tow
bioTowMeta = bioTowTab %>% 
  select(cno, towDate, station, towID, latStart_dd,lonStart_dd,latEnd_dd,lonEnd_dd) %>% 
  distinct() %>% 
  dplyr::mutate(id = 1:n())

bathPath = here("data/environmental/gebco_2026_n41.4_s39.8_w-74.3_e-71.4.nc")
# turn this into a raster object
bathRast = terra::rast(bathPath)
bathDf = as.data.frame(bathRast, xy = TRUE)
bathVars = names(bathDf)[3]
threshold = 0

ggplot()+
  geom_raster(data = bathDf,aes(x = x, y = y, fill = ifelse(.data[[bathVars]] > threshold, NA, .data[[bathVars]])), alpha = 1)+
  annotation_borders('world',
                     xlim = c(NYSbbox[1],NYSbbox[2]),
                     ylim = c(NYSbbox[3], NYSbbox[4]),
                     fill = 'grey')+
  annotate(geom = 'rect', xmin = NYSbbox[1], xmax = NYSbbox[2],
                    ymin = NYSbbox[3], ymax = NYSbbox[4], fill = NA, color = 'black')+
  coord_sf(crs = 4326,
           xlim = c(NYSbbox[1],NYSbbox[2]),
           ylim = c(NYSbbox[3], NYSbbox[4]))+
  # geom_point(data = bioTowTab, aes(x = lonStart_dd, latStart_dd))+
  geom_segment(data = bioTowMeta, aes(x = lonStart_dd, y = latStart_dd, xend = lonEnd_dd, yend = latEnd_dd, group = id))+
  labs(title = "NYS study region", x = 'Longitude', y = 'Latitude')+
  theme(legend.position = 'none')
```

For each of these sampling dates we can extract the environmental data from the collected during the sampling.

```{r get-env}
bioTowMonth = bioTowTab %>% 
  mutate(towmonth = as.Date(paste0(lubridate::year(towDate),"-",lubridate::month(towDate),"-01"), format = "%Y-%m-%d")) %>%
  select(towmonth) %>% 
  distinct %>%
  # join regional sst
  left_join(., readRDS(here('data/environmental/sst.rds')) %>% rename(towmonth = date)) %>% 
  # join regional chla
  left_join(., readRDS(here('data/environmental/chl.rds')) %>% rename(towmonth = date)) %>% 
  # join the regional zoo
  left_join(., readRDS(here('data/environmental/zooc.rds')) %>% rename(towmonth = date)) %>% 
  # join the NAO
  left_join(., readRDS(here('data/environmental/nao.rds')) %>% select(towmonth = date, nao)) %>% 
  # join the AMO 
  left_join(., readRDS(here('data/environmental/amo.rds')) %>% select(towmonth = date, amo = ssta), by = 'towmonth')

## extract the depths for start and end from bathymetry data

bioTowStart = vect(bioTowMeta, geom = c('lonStart_dd','latStart_dd'))
bioTowEnd = vect(bioTowMeta, geom = c('lonEnd_dd','latEnd_dd'))
bioTowMeta$startDepth_m = unlist(terra::extract(bathRast, y = bioTowStart)['elevation'])
bioTowMeta$endDepth_m = unlist(terra::extract(bathRast, y = bioTowEnd)['elevation'])

bioTowMeta = bioTowMeta %>% 
  select(id, cno, towDate, station, towID, lonStart_dd, latStart_dd, lonEnd_dd, latEnd_dd, startDepth_m, endDepth_m) %>% 
  dplyr::mutate(meanDepth_m = (startDepth_m+endDepth_m)/2)
```

## Prepare data for modeling

```{r merge-all-data}
bioEnv = bioTowMonth
    
bioData = bioTowTab %>% 
  dplyr::mutate(towmonth = as.Date( paste0( lubridate::year(towDate), "-", lubridate::month(towDate), "-01"), format = "%Y-%m-%d")) %>%
  dplyr::summarise(abun = n(),
                   biomass = sum(weight),.by = c(towmonth, spn, trawlDist_m))

```


