---
title: "community-analysis"
output: 
  html_document:
    toc: true
    toc_depth: 4
---

<link rel="stylesheet" type="text/css" href="http://cdn.datatables.net/1.10.5/css/jquery.dataTables.min.css">
<script src="http://code.jquery.com/jquery-2.1.2.min.js"></script>
<script src="http://cdn.datatables.net/1.10.5/js/jquery.dataTables.min.js"></script>

<script type="text/javascript">
         $(document).ready(function() {
             $(".dtable").DataTable();
         } );
</script>


```{r init, echo = FALSE, warning=FALSE, message=FALSE}
here::i_am('docs/community-analysis.Rmd')
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  echo = FALSE
)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message= FALSE)

# Define a custom knitr hook for collapsing output
knitr::knit_hooks$set(fig_collapse=function(before, options, envir) {
  if (before) {
    # Opening HTML tags for the collapsible container
    paste('<details><summary>Click to view figures</summary><div>')
  } else {
    # Closing HTML tags
    paste('</div></details>')
  }
})
## load the libraries needed
library(here)
library(junkR)
library(RODBC)
library(magrittr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(rstan)
library(purrr)
library(viridis)
update_data = FALSE
theme_set(theme_minimal())
# source the code to initiate the 
conn <<- odbcConnectAccess2007(here("ignore/Nearshore Survey.accdb"))

# source all the functions to read, clean, and update data products
purrr::walk(list.files(here::here('code/source'), pattern = ".*.R", full.names = TRUE), function(x) source(x))


```

Introduction
------

```{r load-biotow-data}

get_trawl_data(conn = conn, update = update_data)

```

## Data summaries:

How many 'bad' coordinates do we have? Either missing or with minute coordinates greater than 60.

`r bioTowTab %>% dplyr::filter(when_any(is.na(latStart_dd), is.na(lonStart_dd), is.na(latEnd_dd), is.na(lonEnd_dd), latme > 60)) %>% select(cno) %>% unique %>% count`: `r bioTowTab %>% dplyr::filter(when_any(is.na(latStart_dd), is.na(lonStart_dd), is.na(latEnd_dd), is.na(lonEnd_dd), latme > 60)) %>% select(cno) %>% unique %>% knitr::kable(.)`

How many missing weights?

`r bioTowTab %>% dplyr::filter(is.na(weight)) %>% count`

Missing weights with no corresponding length measurements?

`r bioTowTab %>% dplyr::filter(when_all(is.na(weight), is.na(flength), is.na(tlength))) %>% count`

## Length-mass relationships

While directly measured weights are present for most species, some of the data are censored due the precision of the balance, so weights bottom out at 0.01 and, for small species, exhibit distinct 'steps'. This creates an artificial truncation of the lower bound of the data and can also influence biomass estimates. This is likely not a huge deal for the purposes here, but if we want to incorporate this into any analyses, we can model masses and their uncertainty by incorporating the censored data structure directly into a length-mass model.

 
```{r censored-mass-ex}
spList = bioTowTab %>% 
  dplyr::filter(!when_all(is.na(tlength), is.na(flength), is.na(weight))) %>% 
  named_group_split(spn)

spnLWList = spList %>% map(~.x %>% 
    select(tlength, weight) %>% 
    na.omit %>% 
    # create a vector of censored measurements
    mutate(censored = ifelse(weight == 0.01, TRUE,FALSE))) %>% 
    # remove species that have fewer than 10 observations
    keep(~nrow(.) > 10)
# also remove American Lobster as they have a clear bi-modal L-M relationship
    # probably due to sex-specific differences
spnLWList['American Lobster'] <- NULL
    
# using Alewife as an example
spnLWList['Alewife'] %>% 
  map(~.x %>% 
        ggplot()+
        geom_point(aes(x = tlength, y = weight, color = censored))+
        scale_color_viridis(discrete = TRUE, end = 0.8)+
        scale_y_continuous(name = "mass (g)", limits = c(0, NA)))

```

```{r length-mass, warning=FALSE, message = FALSE}
spNames = names(spnLWList)

# debugonce(build_LM)
spNames %>% walk(~build_LM(specName = ., df = bioTowTab))

```

```{r plot length-mass-PPC, fig_collapse = TRUE}
lmFiles = list.files(here('ignore/models/'), pattern = '.*LM_censored.rds', full.names = TRUE)

lmNames = gsub(".*(/\\w+( \\w+)*_LM_censored)\\.rds","\\1", lmFiles) %>% 
  map_chr(~gsub("/(\\w+( \\w+)*)_LM_censored","\\1", .x))

lmFiles= setNames(lmFiles, nm = lmNames)

# debugonce(plot_LM_PPC)
map2(lmFiles, spnLWList, ~plot_LM_PPC(model_path = .x, df = .y))

```

## Collecting environmental variables

Here are the sampling dates from the trawl data.

<details>

```{r trawl-date-info}

bioTowTab %>% 
  mutate(towmonth = as.Date(paste0(lubridate::year(towdate),"-",lubridate::month(towdate),"-01"), format = "%Y-%m-%d")) %>%
  select(towmonth) %>% 
  distinct %>% 
  knitr::kable(.)
```

</details>

For each of these sampling dates we can extract the environmental data collected during the sampling.

```{r }

sqlTables(conn)

```
